### 2020-06-01 case-studies.tex
finished first draft of live evaluation section

To avoid this situation, the livelit expansion mechanism 
maintains parameter and splice hygiene: parameters and splices are included in the expansion 
in a capture-avoiding manner.
Consequently, clients can reason abstractly about binding in splices: variables in splices
always refer to bindings visible to the client, rather than bindings that are hidden inside expansions.
This is implemented by alpha-renaming bindings internal to the expansion as necessary.
(We discuss potentially relaxed variations of this hygiene discipline in Sec.~\ref{sec:discussion}).
\subsubsection{Live Evaluation}\label{sec:live-evaluation}

>>>>

To avoid this situation, the livelit expansion mechanism 
maintains parameter and splice hygiene: parameters and splices are included in the expansion 
in a capture-avoiding manner.
Consequently, clients can reason abstractly about binding in splices: variables in splices
always refer to bindings visible to the client, rather than bindings that are hidden inside expansions.
This is implemented by alpha-renaming bindings internal to the expansion as necessary.
(We discuss potentially relaxed variations of this hygiene discipline in Sec.~\ref{sec:discussion}.)

\subsubsection{Live Evaluation}\label{sec:live-evaluation}
Livelits can request the result of evaluating a splice or a parameter. 
For example, the \li{\$dataframe} livelit uses this facility to display
the evaluation result for each cell when the cursor is not on it, like a spreadsheet.
The \li{\$cutoffs} livelit uses this facility to plot the grades list, provided
as a parameter, on the number line.

\paragraph{Closure Collection} Evaluation in Hazel is, as usual, defined only for closed expressions, 
but parameters and splices can both refer to variables in the surrounding
context. In order to provide evaluation results for parameters or splices containing free variables, 
Hazel performs \textbf{closure collection} in two phases.

In the first phase, Hazel evaluates the program after placing holes in place of each livelit, 
following the evaluation semantics for incomplete programs developed by \citet{HazelnutLive}. 
The result is an expression with holes in it, each of which is equipped with a closure. 
For example, there is a single closure associated with the application of \li{\$dataframe} in Fig.~\ref{fig:grading}.
This closure contains the value of the \li{weights} variable from Line 1 and any other variables in the context 
around that livelit, e.g. 
from imported libraries, not shown. Consequently, when the livelit requests the evaluation result for
splices that use these variables, such as the cell shown selected in Fig.~\ref{fig:grading}, 
an evaluation result is available. (We consider the situation where more than one closure is available
in the Sec.~\ref{sec:image-transformation}.)

Similarly, the closure for the \li{\$cutoffs} livelit in Fig.~\ref{fig:grading} includes a result for 
the \li{grades} variable. However, this variable's result, along with the 
results for other variables that depend on \li{grades}, such as \li{averages}, ultimately depend on the 
expansion of the \li{\$dataframe} livelit. If we stop after the first phase of closure collection,
where these expansions have not yet been included, 
then no useful result will be available here.
For this reason, there is a second phase of closure collection where any livelit holes that appear 
in the collected livelit closures are \emph{resumed}, i.e. the expansion is generated 
(in this case, for the \li{\$dataframe} livelit) and is used to fill 
the hole before evaluation resumes, following the semantics for resumption developed by \citet{HazelnutLive}.
Livelit expansions do not themselves contain livelits, so no subsequent resumption phases are necessary.

The final evaluation result for the program as a whole can also proceed by resumption, because 
closure collection performs all of the computations that do not ultimately depend on the livelit expansion.
Consequently, Hazel does not need to evaluate the program multiple times in order to support edit-time 
(i.e. live) evaluation in this way.

\paragraph{Indeterminate Results}
We used the phrase \emph{evaluation result} rather than \emph{value} above purposefully: 
splices and parameters can be incomplete, i.e. they can contain holes, so 
evaluation does not always result in a value \cite{HazelnutLive}.
In particular, when a hole appears in elimination position, e.g. in function position of a function application,
 the result is instead an indeterminate expression. 
When a livelit requests an evaluation result, it must be able to handle indeterminate results.
For example, if the parameter to \li{\$cutoffs} were a hole or some other indeterminate expression\todo{small screenshot?}{}, 
then it would have degraded functionality 
(in this case, it would display the list elements that are values on the timeline, skipping indeterminate elements\todo{implement this}{}). 
We will return to how livelit implementations handle this situation in Sec.~\ref{sec:implementation}.
\todo{make sure we cover this}{}
